A Wide-Coverage Symbolic Natural Language Inference System

### 2019-07-15 paper2018/paper-long.tex
paper long version

\documentclass[11pt]{article}
\usepackage{nodalida2019}
\usepackage{times}
\usepackage{todonotes}
\usepackage{tabularx}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

@@

In this paper, we concentrate on this sort of fine-grained, logical
reasoning. In particular, we present a logic-based system that deals
with many linguistic phenomena \emph{at the same time}. It is the
first systems covering the sections on ellipsis and anaphora in the
FraCaS test suite and has the best coverage and accuracy on the
overall test suite.
% This is listed below; we can let the reader figure it out.
% \begin{itemize}
% \item A two-phase system (rather than a one-phase as described by
%   \citet{bernardy_type_2017}): a) using Haskell as a converter between
%   Grammatical Framework trees and Coq types, b) using various type
%   combinators to reach the desired NL semantics.
% \item The first run of the Anaphora and Ellipsis sections of the
%   FraCaS (based on work by \citet{bernardy_jolli})
%  \item  More coverage than any previous logical approach
%  \item State-of-the-art results in terms of accuracy compared to all
%    previous approaches, for every section of the FraCaS
%   \item Handing of multiple readings
%   \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
%   \item New handling of prepositions and adverbs compared to the
%     earlier system in \citet{bernardy_type_2017}, as well as new
%     handling of definitions, enabled by the dynamic semantics system.
%   \item Improvement over the previous treatment of Plurals/Quantifiers
%   \item Big improvement on the comparatives section enabled by
%     handling comparatives using a threshold (BUT one different to the
%     one used for adjectives). Again, the dynamic semantics system is
%     crucial here.
% \end{itemize}

@@

\paragraph{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP \to CN \to CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete syntax of various natural languages.
% These mappings can be inverted by GF, thus offering parsing from natural text into abstract syntax.
However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity.

@@

% The architecture is represented schematically in Figure \ref{fig:overview}.
% \begin{figure*}
%   \centering
% {
%   \tiny
% \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
%                     semithick]
%   \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]

%   \node[state]         (txt)                {text};
%   \node[state]         (ast) [right of=txt] {AST};
%   \node[state]         (pp)  [right of=ast] {Probabilistic program};
%   \node[state]         (res) [right of=pp] {result in $(0,1)$};

%   \path (txt) edge              node (GF)  {GF}                  (ast)
%         (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
%         (pp) edge               node (Coq) {Coq} (res);

%   \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
%   \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
%   \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};

%   \path
%   (Syntax) edge (GF)
%   (Sem) edge (Hask)
%   (Hyper) edge (Coq);
% \end{tikzpicture}
% }
% \caption{Phases in our system. ($\ast$) At the level of Coq, we handle
%   the details of the adverbial (veridicality properties) and
%   adjectival semantics (division into subsective, extentional,
%   non-committal, etc. categories.) }
%   \label{fig:overview}
% \end{figure*}

@@

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost $\theta$ orders. $\theta$ $\ge$ 0.  ITEL won at least $\theta$+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions.\footnote{The degree parameter assumption is not new in the formal semantics literature \cite{Cresswell:1976,Heim:2001,Kennedy:2007,chatzikyriakidis:2017} among many others. The specific details and computational implementation, however, are.}
% SC
%Finally, ``more'' is systematically intepreted as ``at least
%<elliptic quantity>+1''. This treatment, which we illustrated here on
%an example, is systematic in our implementation.

@@

% \paragraph{Adverbs}
% Another point, of minor theoretical importance but major practical
% one, is our handling of adverbial phrases. We interpret adverbs (and
% in general all adverbial and prepositional phrases) as VP-modifiers:
% $Adv = VP \to VP$, where $VP = object \to Prop$. However, applying adverbs
% to verb-phrases heavily complicates the Coq proofs, because such
% phrases can contain quantifiers. Therefore, we instead move the
% adverbs, so that they apply to (atomic) verbs only. Proofs can then be
% simplify accordingly. \todo{theoretical justification in the
%   literature?}

@@

We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
total of 259 cases. We excluded only section 7, ``temporal reference''.
% The reason for
% doing so is that, in our view, it contains too many examples which
% require \textit{ad-hoc} treatment, and thus makes little sense to
% include without complementing it with a more thorough suite which
% captures a more complete landscape of the phenomena that section 7
% touches.

FraCaS classifies each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by \citet{maccartney:2007} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labellings (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of 6 cases
%(SC: Shall we remove this to the supplementary material and just say to have a look at how we amended the suite?)
% the cases listed in Table \ref{tab:overrulings},
which appeared to have been
misclassified. We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized. In terms of comparison, this only has a minor effect, since our system is the first system to run sections 3 and 4. 
% \begin{table}
%   \centering
%   \small
% \begin{tabularx}{\columnwidth}{ccp{3.8cm}}
%  test case & new class & comment \\ \hline
%     005 & UNK & missing hypothesis: there are italian tenors \\
%    056 &  Yes & Already identified as such by MacCartney \\
%   069 & Unk & Mary could have used someone else's workstation \\
%   119 & Unk & \textit{ibid.} \\
%   181 & Yes & for the same reason as 180 \\
%   226 & Yes &
% \end{tabularx}
%   \caption{Overruled FraCaS cases}
%   \label{tab:overrulings}
% \end{table}

>>>>

\documentclass[11pt]{article}
% \usepackage{nodalida2019}
\usepackage{natbib}
\usepackage{times}
\usepackage{todonotes}
\usepackage{tabularx}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

@@

In this paper, we concentrate on this sort of fine-grained, logical
reasoning. In particular, we present a logic-based system that deals
with many linguistic phenomena \emph{at the same time}. It is the
first system covering the sections on ellipsis and anaphora in the
FraCaS test suite and has the best coverage and accuracy on the
overall test suite.
% This is listed below; we can let the reader figure it out.

In sum our contributions are as follows:
\begin{itemize}
\item A two-phase system (rather than a one-phase as described by
  \citet{bernardy_type_2017}): a) using Haskell as a converter between
  Grammatical Framework trees and Coq types, b) using various type
  combinators to reach the desired NL semantics.
\item The first run of the Anaphora and Ellipsis sections of the
  FraCaS (based on work by \citet{bernardy_jolli})
 \item  More coverage than any previous logical approach
 \item State-of-the-art results in terms of accuracy compared to all
   previous approaches, for every section of the FraCaS
  \item Handing of multiple readings
  \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
  \item New handling of prepositions and adverbs compared to the
    earlier system in \citet{bernardy_type_2017}, as well as new
    handling of definitions, enabled by the dynamic semantics system.
  \item Improvement over the previous treatment of Plurals/Quantifiers
  \item Big improvement on the comparatives section enabled by
    handling comparatives using a threshold (BUT one different to the
    one used for adjectives). Again, the dynamic semantics system is
    crucial here.
\end{itemize}

@@

\paragraph{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP \to CN \to CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete syntax of various natural languages.
These mappings can be inverted by GF, thus offering parsing from natural text into abstract syntax.
However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity.

@@

The architecture is represented schematically in Figure \ref{fig:overview}.
\begin{figure*}
  \centering
{
  \tiny
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
                    semithick]
  \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]

  \node[state]         (txt)                {text};
  \node[state]         (ast) [right of=txt] {AST};
  \node[state]         (pp)  [right of=ast] {Probabilistic program};
  \node[state]         (res) [right of=pp] {result in $(0,1)$};

  \path (txt) edge              node (GF)  {GF}                  (ast)
        (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
        (pp) edge               node (Coq) {Coq} (res);

  \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
  \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
  \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};

  \path
  (Syntax) edge (GF)
  (Sem) edge (Hask)
  (Hyper) edge (Coq);
\end{tikzpicture}
}
\caption{Phases in our system. ($\ast$) At the level of Coq, we handle
  the details of the adverbial (veridicality properties) and
  adjectival semantics (division into subsective, extentional,
  non-committal, etc. categories.) }
  \label{fig:overview}
\end{figure*}

@@

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost $\theta$ orders. $\theta$ $\ge$ 0.  ITEL won at least $\theta$+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions.\footnote{The degree parameter assumption is not new in the formal semantics literature \cite{Cresswell:1976,Heim:2001,Kennedy:2007,chatzikyriakidis:2017} among many others. The specific details and computational implementation, however, are.}
%
Finally, ``more'' is systematically intepreted as ``at least
<elliptic quantity>+1''. This treatment, which we illustrated here on
an example, is systematic in our implementation.

@@

\paragraph{Adverbs}
Another point, of minor theoretical importance but major practical
one, is our handling of adverbial phrases. We interpret adverbs (and
in general all adverbial and prepositional phrases) as VP-modifiers:
$Adv = VP \to VP$, where $VP = object \to Prop$. However, applying adverbs
to verb-phrases heavily complicates the Coq proofs, because such
phrases can contain quantifiers. Therefore, we instead move the
adverbs, so that they apply to (atomic) verbs only. Proofs can then be
simplify accordingly.

@@

We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
total of 259 cases. We excluded only section 7, ``temporal reference''.
The reason for
doing so is that, in our view, it contains too many examples which
require \textit{ad-hoc} treatment, and thus makes little sense to
include without complementing it with a more thorough suite which
captures a more complete landscape of the phenomena that section 7
touches.

FraCaS classifies each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by \citet{maccartney:2007} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labellings (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of 6 cases
%(SC: Shall we remove this to the supplementary material and just say to have a look at how we amended the suite?)
% the cases listed in Table \ref{tab:overrulings},
which appeared to have been
misclassified. We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized. In terms of comparison, this only has a minor effect, since our system is the first system to run sections 3 and 4. 
\begin{table}
  \centering
  \small
\begin{tabularx}{\columnwidth}{ccp{3.8cm}}
 test case & new class & comment \\ \hline
    005 & UNK & missing hypothesis: there are italian tenors \\
   056 &  Yes & Already identified as such by MacCartney \\
  069 & Unk & Mary could have used someone else's workstation \\
  119 & Unk & \textit{ibid.} \\
  181 & Yes & for the same reason as 180 \\
  226 & Yes &
\end{tabularx}
  \caption{Overruled FraCaS cases}
  \label{tab:overrulings}
\end{table}
